---
title: "Bee Data"
output: 
  html_document: 
    keep_md: yes
---
##Questions:
1. There’s a dataset in the folder called, “Bee colony Census Data by County”  and we wanted help in merging the data sets. 


2. Also in the “Bee colony Census Data by County” data set, there’s a column named “value” and we’ve looked in the glossary but we couldn’t find any info on it. We were wondering if you knew what we could use the “value” column for or if it’s even worth using anything in the ”Bee colony Census Data by County” data set. 

3. I'm having trouble geting all the columns in the object "bee2" Renamed. It seems like it won't recognize the variables original names even before I've renamed them. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#install.packages("tidyverse")
#install.packages("skimr")
#install.packages("RColorBrewer")
library("RColorBrewer")
library(tidyverse)
library(skimr)
```


```{r}
#if you guys are having trouble with viewing it, move the data to the same folder. ie move the rmd file to the finley data folder or move the excel sheet to the "animal-master" folder
bee <- 
  readr::read_csv("bee2.csv")
bee
names(bee)
```

```{r}

bee <- 
  bee %>% 
  dplyr::rename(
          year        = "Year",
          season      = "Season",
          state       = "State",
          totAnnualLoss = "Total Annual Loss",
          beekeepers = "Beekeepers",
          beekeepersExclusive = "Beekeepers Exclusive to State",
          colonies = "Colonies",
          coloniesExclusive =  "Colonies Exclusive to State" 
          )

names(bee)
```


```{r}
#Be keeper vs Bee loss
bee %>% 
  filter(year=="2016/17") %>% 
  ggplot(aes(x=state, y= beekeepers, fill = state))+
  geom_col()+
  #coord_flip()+
  labs(title = "Number of Beekeepers per State in 2016-2017",
       x = "State",
       y = "Number of Beekeepers Hired by State")+
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=10))+
  
  #scale_fill_brewer(palette = "Dark2")
  
  
```

```{r}
ggplot(data=bee2, mapping=aes(x="Year" , y=" Total Annual Loss")) +
  geom_jitter()+
  geom_smooth(method=lm, se=FALSE)
```

```{r}
ggplot(data=bee2, mapping=aes(x=" State", y=" Beekeepers"))+
  geom_boxplot()
```
```{r}
names(bee2)

```
```{r}
names(bee1)

```







