---
title: "Bees!"
author: "Isha, Jade, Valerie"
date: "Winter 2019"
output: 
  ioslides_presentation: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#library("jpeg")
```

## Packages
We use the tidyverse and skimr for our project.
```{r echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
#install.packages("skimr")
#install.packages("tidyverse")
library(tidyverse)
library(skimr)
library("RColorBrewer")
```

## Load the data
We found our data at the https://data.world/datasets/bees
```{r message=FALSE, warning=FALSE, include=FALSE}
bee <- 
  readr::read_csv("/Users/clmuser/Desktop/Animals-master/Animals-master 5/bee2.csv")
bee
```

##Rename the Data
```{r}
bee <- 
  bee %>% 
  dplyr::rename(
          year        = "Year",
          season      = "Season",
          state       = "State",
          totAnnualLoss = "Total Annual Loss",
          beekeepers = "Beekeepers",
          beekeepersExclusive = "Beekeepers Exclusive to State",
          colonies = "Colonies",
          coloniesExclusive =  "Colonies Exclusive to State" 
          )

names(bee)
```

```{r}
options(scipen=999)
```

## How many bee colonies are there in each state?
```{r echo=FALSE, message=FALSE, warning=FALSE}
bee %>%
  filter(year=="2010/11") %>%
ggplot( aes(x=state,y=colonies, fill= state)) +
  geom_bar(stat = "identity") +
  labs(title = "Colonies Vs State (2010/2011)",
       x = "states",
       y = "colonies")+
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=10),
          legend.position="none",
          plot.title = element_text(lineheight=.8, face="bold"))
  scale_fill_brewer(palette = "Dark2")
```

## Colonies in 2016/2017
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
bee %>%
  filter(year=="2016/17") %>%
  ggplot( aes(x=state,y=colonies, fill = state)) +
  geom_bar(stat = "identity") +
  labs(title = "Colonies Vs State (2016/2017)",
       x = "states",
       y = "colonies")+
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=10),
          legend.position="none",
          plot.title = element_text(lineheight=.8, face="bold"))
  scale_fill_brewer(palette = "Dark2")
```

## Fun Facts
- California has a discrepancy as they started a new initive to revive the Honey bee population by having them pollinate almonds. Each almond acre requires two honey hives

##Is it true bee colonies are losing?
```{r}
bee_2010<- 
  bee %>%
  filter(year =="2010/11") %>% 
  select(year, state, totAnnualLoss) %>% 
  arrange(desc(totAnnualLoss))
bee_2010
```

## Top 5 states with the highest colony loss in 2010
```{r}
bee_2010 %>% 
  filter(totAnnualLoss >=0.75) %>% 
  ggplot(aes(state, totAnnualLoss, fill = state))+
  geom_col()+
  labs(title= "Top 5 States with the Highest Colony Loss in 2010")+
  theme(axis.text.x = element_text(hjust= 0.95, size=rel(1)))+
  theme(plot.title = element_text(size = rel(1), hjust = 0.5))+
  theme_bw()
```

##In the most recent year 2016-2017
```{r}
bee_2016<-
  bee %>% 
  filter(year =="2016/17") %>% 
  select(year, state, totAnnualLoss) %>% 
  arrange(desc(totAnnualLoss))
bee_2016
```

##Top 5 states with the highest colony loss in 2016
```{r}
bee_2016 %>% 
  filter(totAnnualLoss >=0.65) %>% 
  ggplot(aes(state, totAnnualLoss, fill = state))+
  geom_col()+
  labs(title= "Top 5 States with the Highest Colony Loss in 2017")+
  theme(axis.text.x = element_text(hjust= 0.95, size=rel(1)))+
  theme(plot.title = element_text(size = rel(1), hjust = 0.5))
```


##Average lossage of the colony in U.S.
```{r}
average_loss<-
bee %>% 
  group_by(year) %>% 
  summarize(mean_annualLoss = mean(totAnnualLoss,na.rm = TRUE))
average_loss
```

##Average Annual loss through out the year
```{r}
average_loss %>% 
  ggplot(aes(year,mean_annualLoss,size = mean_annualLoss, color=year))+
  geom_point()+
  labs(title = "Average Colony Loss from 2010 to 2017",
       x = "Year",
       y = "Mean Annual Loss")+
  theme(axis.title = element_text(hjust = 0.95))+
  theme(axis.text.x = element_text(size= rel(1)))
```


##BUT WHY ARE BEE COLONIES DESCREASING?
##Let's investigate one posible reason: Beekeepers

##Here's How Many Beekeepers Were Employed by Each State in 2016-2017
```{r}
bee %>% 
  filter(year=="2016/17") %>% 
  ggplot(aes(x=state, y= beekeepers, fill = state))+
  geom_col()+
  labs(title = "Number of Beekeepers per State in 2016-2017",
       x = "State",
       y = "Number of Beekeepers Hired by State")+
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=10),
          legend.position="none",
          plot.title = element_text(lineheight=.8, face="bold"))
```

##Pennsylvania and Virginia Are states with higher bee loss 
- Pennsylviania is ranked #16 in highest total annual bee loss
- Virginia is ranked #19

##There is no relationship between number of colonies and beekeepers
```{r}
bee %>% 
  filter(year=="2016/17") %>% 
  ggplot(aes(x=beekeepers, y=totAnnualLoss, color = state))+
  geom_jitter()+
  geom_smooth(method=lm, se=FALSE)+
  labs(title = "Total Annual Loss vs Number of Beekeepers in 2016-2017",
       x = "Number of Beekeepers Hired by State",
       y = "Total Annual Loss of Bee Colonies")+
   theme( axis.text.x = element_text(angle = 60, hjust=0.8),
          axis.text=element_text(size=8),
          axis.title=element_text(size=10),
          legend.position="none",
          plot.title = element_text(lineheight=0.8, face="bold"))+
  scale_fill_brewer(palette = "Dark2")

```

##What really matters is the ratio of beekeepers to colonies, Right?

##But...
```{r}
bee %>% 
  mutate(coloniesPerKeeper = colonies/beekeepers) %>% 
  filter(year=="2016/17") %>% 
  ggplot(aes(x=coloniesPerKeeper, y= totAnnualLoss, color=state))+
  geom_jitter(shape=18, alpha=.8)+
  geom_smooth(method=lm, se=FALSE)+
  labs(title = " Ratio of Colonies to Beekeepers vs Total Annual Loss in 2016-2017",
       x = "Total Annual Loss",
       y = "Colonies  /  Beekeepers")+
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text.y = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=10),
          legend.position="none",
          plot.title = element_text(lineheight=.8, face="bold"))

```

##Furthermore...
```{r}
bee %>% 
  mutate(coloniesPerKeeper = colonies/beekeepers) %>% 
  filter(state == "Oklahoma" | state=="Iowa") %>% 
  ggplot(aes(x=totAnnualLoss, y= coloniesPerKeeper, color = year))+
  geom_jitter()+
  geom_smooth(method=lm, se=FALSE)+
  labs(title = "Total Anuual Loss vs Ratio of Colonies to Beekeepers",
       x = "Total Annual Loss",
       y = "Colonies : Beekeepers")+ 
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text.y = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=13),
          plot.title = element_text(lineheight=.8, face="bold"))+
  facet_grid(. ~ state)+
  scale_fill_brewer(palette = "OrRd")
```

-Iowa and Oklahoma are two states with some of the highest total annual loss.

##On the other hand...
```{r}
bee %>% 
  mutate(coloniesPerKeeper = colonies/beekeepers) %>% 
  filter(state == "California" | state== "North Dakota") %>% 
  ggplot(aes(x=totAnnualLoss, y= coloniesPerKeeper, color = year))+
  geom_jitter()+
  geom_smooth(method=lm, se=FALSE)+
  labs(title = "Total Anuual Loss vs Ratio of Colonies to Beekeepers",
       x = "Total Annual Loss",
       y = "Colonies : Beekeepers")+ 
   theme( axis.text.x = element_text(angle = 60, hjust=1),
          axis.text.y = element_text(angle = 60, hjust=1),
          axis.text=element_text(size=8),
          axis.title=element_text(size=13),
          plot.title = element_text(lineheight=.8, face="bold"))+
  facet_grid(. ~ state)+
  scale_fill_brewer(palette = "BuGn")
```

-California and North Dakota typically have a low total annual loss. 

-We can see that once again, there is no linear regression line


##Conclusion

The number of beekeepers being hired is not related to rising rates colony collapse. But here are some other factors that are likely to be:

- Pesticides, fungicides, and insecticides
- Bees' diets are not exactly healthy, but often consist of junk food in the form of sugar or corn syrup
- erratic weather patterns brought on by a changing climate

Source: https://www.motherjones.com/environment/2015/07/climate-change-killing-bumblebees/
http://sfyl.ifas.ufl.edu/agriculture/colony-collapse-disorder/

